\documentclass{scrartcl}

\usepackage{biblatex}
\bibliography{refs}

\usepackage{tikz-cd}

\usepackage{amssymb}

\usepackage{hyperref}

\title{Category Theory Study Group}
\date{}

\begin{document}

\maketitle

\section*{Format}
Denis Redozubov asked me to help organize a study group for category theory.
As format for this study group I propose that we read about a
particular concept in a category theory book or on \url{ncatlab.org}.
In addition, I will post exercises about the topics and we discuss our
solutions in a group session. But ultimately I would like this group
to become self-sustained:

Category theory introduces many abstractions that were derived from
properties of sets and functions.
A good way to learn what these abstractions mean is to look at
instances of these abstractions in different settings (i.e. in
different categories).
An even better way to learn is to work out these instances yourself in
a setting you are familiar with.
I encourage you to pick a category that is of interest to you
(e.g. the category of database tables as objects and foreign keys as
arrows) and work out what the structure of this category is as we go
through the book, i.e. what are functors from and into this category,
what are initial and terminal objects, etc.
If you come up with good examples, please share them with the group
that everyone can learn from them.

\section*{First Session}
For the first session, read 1.1--1.5 of \citetitle{Awodey2010} by
\citeauthor{Awodey2010}~\cite{Awodey2010}. In addition you might want
to watch recordings of the category theory lectures of Steve
Awodey\footnote{\url{https://www.cs.uoregon.edu/research/summerschool/summer12/curriculum.html}}
or the lectures of Bartosz Milewski\footnote{\url{https://youtu.be/I8LbkfSSR58}}.

\newpage

\section*{Exercises}

\begin{enumerate}

\item
  Show that database schemas can be regarded as categories, where
  objects are tables and the arrows are foreign keys.  What are the
  identities and composition operations?  Try to find examples of
  categories in your own research area/work.

\item \label{ex:matrix} Arrows in a category are not necessarily
  functions.  Define a category where the objects are positive numbers
  and an arrow from $n \rightarrow m$ is a $m \times n$ matrix.  What
  is a suitable composition operation?

\item
  A diagram commutes if all arrows with the same domain and codomain
  are equal.  What does it mean that the following diagram commutes?
  Write down all equalities between pairs of arrows and their
  compositions that hold.
  \begin{center}
  \begin{tikzcd}
    A \arrow[r, "f_1", bend left]\arrow[r, "f_2", bend right] \arrow[d,"f_3"] & B \arrow[d,"f_4"] \\
    C \arrow[r, "f_5", bend left]\arrow[r, "f_6", bend right] & D
  \end{tikzcd}
  \end{center}

\item
  Partial orders are categories where objects are elements and there
  is exactly one arrow between objects $a$ and $b$ iff $a \leq b$.
  Draw a diagram of the category generated by the partial order of the
  power set of $\{1,2,3\}$ and subset inclusion.  What are the
  identities of this categories and why does there exists a
  composition?  Prove that a functor between two partial orders
  regarded as categories is a monotone function.

\item
  Monoids are sets/types equipped with an distinguished element called
  unit and an associative binary operation.
  %\footnote{Haskell includes a comprehensive library for monoids: \url{https://hackage.haskell.org/package/base/docs/Data-Monoid.html}. You might want to study the monoids $\mathtt{Any}$, $\mathtt{All}$, $\mathtt{First}$, $\mathtt{Last}$, $\mathtt{Max}$, $\mathtt{Min}$, $\mathtt{Sum}$, $\mathtt{Product}$,  $\mathtt{Endo}$, $\mathtt{Dual}$, for tuples and for functions.}
  Show that lists under the append operation form a monoid.
  Show that the $\mathtt{Option}/\mathtt{Maybe}$ type is a monoid.
  Show that the natural numbers under addition are a monoid.

  We now consider the category of monoids and their homomorphism.
  Come up with a commuting diagram between the three monoids from
  before.  In other words, find at least three monoid homomorphisms
  between these types and show that all arrows with the same domain
  and codomain are equal.

\item
  The image of functors can be used to describe a specific diagram in
  a category.  Define a functor with a suitable domain, whose image is
  the diagram of the previous exercise.  Define a functor with a
  suitable domain whose image is this diagram.
  \begin{center}
  \begin{tikzcd}[row sep=large, column sep=small]
     & A \arrow[rd,"f_3"] & \\
    B \arrow[ru, "f_2"] \arrow[rr,"f_6"] & & C \arrow[d,"f_4"] \\
    D \arrow[rr,"f_5"] \arrow[u,"f_1"]  & & E
  \end{tikzcd}
  \end{center}
  How do the images of other functors with the same domain as the
  previous functor look like?

\item Functors transfer commuting diagrams in one category to another
  category.  To this end, prove if the upper face of the following
  diagram commutes, so does the lower face.
  \begin{center}
  \begin{tikzcd}[row sep=small]
    &  & B \arrow[rd,"g"] & \\
    \mathbf{C} \arrow[ddd,"F"] & A \arrow[ru, "f"]\arrow[rd, "i"{below left}]  & & D \\
    &  & C \arrow[ru,"j"{below right}] & \\
    &  & FB \arrow[rd,"F g"] & \\
    \mathbf{D} & FA \arrow[ru, "F f"]\arrow[rd, "F i"{below left}] & & FD \\
    &  & FC \arrow[ru,"F j"{below right}] &
  \end{tikzcd}
  \end{center}

\item
  Show that there is a category $\mathbf{Cat}$, whose objects are
  categories and arrows are functors.  What are identity functors?
  How is the composition of two functors defined?  Prove the identity
  laws and that composition is associative.

\item
  An isomorphism between objects indicates, that the objects might
  have different structure, however, they behave the same way.  From a
  perspective of a category, these objects are indistinguishable as
  there is the same number of arrows pointing into these objects and
  the same number pointing away from these objects, as we shall see
  later.  For category theorists, this is a notion of equivalence of
  objects that is fine-grained enough and it is often not worth it to
  study differences between isomorphic objects.  In this vein, prove
  that the relation $X \cong Y$ is an equivalence relation.

  In $\mathbf{Set}$ isomorphisms are bijective functions.  In
  $\mathbf{Mon}$ isomorphisms are bijective monoid homomorphisms.
  However, in other categories, isomorphisms might be something
  completely different.  What are isomorphisms in the category of
  matrices of exercise \ref{ex:matrix}?
\end{enumerate}

\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
